Theories of deliberated preference

### 2018-02-25 simple.tex
Nicer pres and intro

\begin{document}
\title{Simple models of deliberated preference}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

\section{Motivation}
We define the deliberated preference of an individual. We are interested in a context of a choice of a subset of alternatives from a set of possible alternatives given a priori. We define a model of the DP and give conditions and a validation procedure that ensure that what the model claims is in the DP is indeed. We want in particular to allow for the model to be applicable to various individuals that start with different knowledge: the model may use different counter-arguments to convince different individuals.

More in detail. A full-blown proof may be so long that nobody would read it. (That’s why mathematicians do not write their proofs in set-theoretic bare notation.) Hence, shortcuts are taken, which rely on the individual knowing stuff already. More generally, assume you want to convince someone of fact $F$ with a text (we call this an argument), in our case, we want to convince a person that such alternative from the set is a good choice for her. The argument may assume the reader knows already why such or such possible counter-arguments are invalid, and the argument will thus not delve into the details of some points. When shown in interactive mode (our context here), an individual may however question some points (which we call a counter-argument). The argumenter should then be able to provide an answer to the query (a counter-counter-argument).

This article aims at defining such a procedure and conditions of its validity.

\section{Decision situation}
Here is the definition of a decision situation.
\begin{itemize}
	\item All alternatives $\allalts$.
	\item Topic $\alltopic$ = $\Set{\prop_\alt, \alt \in \allalts} ∪ \Set{\prop_{¬\alt}, \alt \in \allalts}$. Denoted simply $\alt$ and $¬\alt$. We define $¬\prop$, with $\prop = \prop_\alt$, as equal to $\prop_{¬\alt}$ and $¬\prop$, with $\prop = \prop_{¬\alt}$, as equal to $\prop_{\alt}$.
	\item All possible arguments: $\allargs$. It contains at least $\emptyset$, the empty argument. It contains all the arguments and more. As an example, $\allargs$ could be the set of all strings (assuming all arguments can be transmitted textually).
\end{itemize}

Define $f_\prop$ from any unordered pair of arguments $(\ar, \ar')$ into $\powerset{\prop\text{-sure}, ¬\prop\text{-sure}, b)}$ (where $b$ means both possible). We present both arguments to $i$ and ask which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind. The set designates all cumulated answers over time, thus allowing $i$ to be unstable (because she has learnt new things in between two questioning, or just because of time passing, or for any known or unknown reason). We assume $b$ is observable as well.
We can also use $\emptyset$ as one or both of the arguments.

@@

@@

TODO.
When “both” is not observable, only the sets \{$\prop$-sure, both\} and \{$¬\prop$-sure, both\} can be observed, and there’s only $\phi=\prop$ and $¬\phi=¬\prop$. Accepting $\prop$ requires an arg that always leads to \{$\prop$-sure, both\}.

The deliberated judgment of $i$ maps any pair $(\prop, ¬\prop)$ to one of four possibilities: $\prop$ is sure, $¬\prop$ is sure, both are possible, or undecided. $\prop$ is sure iff $\exists \ar \ \propsuredec$; both are possible iff $\exists \ar \ \proppossdec$ and $\exists \ar \ ¬\proppossdec$; and undecided is when none of the other three cases hold.
For easier treatment, we summarize $i$’s deliberated judgment in a single set $\iPhi \subseteq \Phi$, as follows: $(\prop, \text{possible}) \in \iPhi ∧ (¬\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $(\prop, ¬\prop)$ maps to both possible; $(\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \in \iPhi ∧ (¬\prop, \text{possible}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $(\prop, ¬\prop)$ maps to $\prop$ is sure, $\phi \notin \iPhi$ for all $\phi$ about $\prop$ iff $(\prop, ¬\prop)$ maps to undecided.
\footnote{Example of $(\prop, ¬\prop)\text{-undecided}$: $\forall (\ar, \ar'), (\ar, \ar') \ileadstose ¬\prop, (\ar, \ar') \ileadstose \prop$.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}

Observe that (thanks to A2) $(\prop, \text{possible}) \in \iPhi$ iff $\exists \ar \ \proppossdec ∧ (\exists \ar \ \propsuredec ∨ \exists \ar \ ¬\proppossdec)$.

\begin{proposition}
	a) $\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$. b) $(\prop, ¬\prop)$ maps to undecided $⇔ \phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$. c) If $\phi \in \iPhi$ for some $\phi$ about $\prop$, then $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$ for all $\phi$ about $\prop$.
\end{proposition}
\begin{proof}
	a) is because some $\ar^+$ is then $\phidec$, thus given any $\ar^−$, $(\ar^+, \ar^−) \ileadstost \phi$ hence $(\ar^+, \ar^−) \ileadstoe \phi$ (using \cref{ft:sttoe}).
\end{proof}

\section{Theories proposing decisive argument}
In most situations, there are too many arguments to exhaustively test whether any $\ar$ is a decisive argument for any $\phi$. Furthermore, it is impossible to confirm definitely that $(\ar, \ar') \ileadstost \phi$, whenever $(\ar, \ar') \ileadstost \phi$. We rather propose to introduce theories, and expose them to falsification.

@@

As is usual in science, falsification procedures should not be considered as ways of definitely validating a theory, but rather as ways of confronting theories.
We say that two theories $T, T'$ disagree on $\phi$ (or equivalently, on $¬\phi$) iff $\phi \in \Phi_{d, T}$ and $¬\phi \in \Phi_{d, T'}$ in some situation $d \in D_T ∪ D_{T'}$. We say that $T$ and $T'$ disagree iff they disagree on some $d \in D_T ∪ D_{T'}$ and $\phi \in \Phi$.
To confront a theory $T$ to a disagreeing theory $T'$, suffice to test, for any $\phi$ such that $\phi \in \Phi_T, ¬\phi \in \Phi_{T'}$, whether $(f_T(\phi), f_{T'}(¬\phi)) \ileadstoe \phi$.
Observe that such a test necessarily falsifies one of the theories. Observe also that this applies to a single theory $T$, that disagrees with itself, if $\Phi_{d, T}$ contains both $\phi$ and $¬\phi$ for some $d$ and some $\phi$.

The following proposition shows that this notion of falsification is sufficient to test the claims of the theory.
\begin{proposition}
	A theory $T$ proposing decisive arguments has no falsification instances iff $\forall d \in D_T, \phi \in \Phi_{d, T}: f_{d, T}(\phi)$ is $\phidec$. Furthermore, if a theory $T$ proposing decisive arguments has no falsification instances, then $\Phi_{d, T} \subseteq \iPhi$, $\forall d \in D_T$.
\end{proposition}

@@

%\bibliography{mybib}

>>>>

\begin{document}
\title{Theories of deliberated preference}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

\begin{abstract}
	In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, as famously written by \citeauthor{von_neumann_theory_2004}. This conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. This article introduces, and defines formally, the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments. As defined, deliberated preferences are only indirectly observable. To make it possible to model deliberated preferences, the article also defines formally the notion of a theory of deliberated preferences, together with a protocol of falsification of such a theory. The proposed falsification protocol involves only observations similar to those used in classical theories of revealed preferences. It is shown that a theory of deliberated preference captures effectively a known subset of the deliberated preferences if and only if it is not falsifiable.
\end{abstract}

\section{Motivation}
In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, following the terms used by \citet{von_neumann_theory_2004} (vNM) \citep{fishburn_retrospective_1989, cailloux_reasons_2017}. Though appropriate when studying the everyday behavior of individuals, this conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. In this article I define the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments.

After having defined this concept formally, I define the notion of a theory pretending to capture the deliberated preferences of a set of individuals about some topic, I propose an empirical procedure to attempt to falsify such a theory, and I show that such a theory tells the truth whenever it is not falsifiable by that procedure.

This article does not attempt to propose a concrete theory of deliberated preferences. It is more preliminarly. It proposes formal definitions for 1) the kind of object that are to be studied for capturing “more reflexive” preferences than considered usually, 2) what theories of such preferences should look like, 3) how to falsify such theories and using which information. Therefore, it proposes a research program rather than a direct solution to the problem of capturing reflexive preferences.

Before delving into the formalism, let me present a bit of motivation for considering deliberated preferences as an interesting object to study.
Being able to capture deliberated preferences can help provide a better understanding of human preferences in the large sense and may improve policy making (though this article does not explore issues of group decision). Besides this very broad goal, here are three example of contexts in which this notion is useful, which may give it a more concrete interest and serve to introduce this concept informally by indicating the kinds of things that it should be able to help understand.

First, decision theory has produced several models of how an individual ought to decide, such as the seminal one of vNM. (The following remarks apply equally well to other normative models such as the one of \citet{savage_foundations_1972}, \citet{keeney_decisions_1993} or more recent ones.) Those models can be used to help individuals take “good” decisions, as opposed to predict their behavior (for which other models, taking into account cognitive limitations, are more appropriate \citep{wakker_prospect_2010}). But it is unclear how to validate recommendations obtained on the basis of such models. Some researchers have opposed the norms on which some of those models are based \citep{ellsberg_risk_1961, allais_so-called_1979}. As \citet{raiffa_back_1985} points out, while it is the case that some sophisticated decision makers may consciously refuse to abide by the vNM axioms (or others), it is also true that some decision makers simply fail to comply with them because of insufficient reflexion. However, no general proposal, as far as I know, has been described yet to test whether, for a given individual and context, those axioms would, after careful thinking, reveal acceptable. Psychologists have tried to test the convincing power of those axioms \citep{slovic_who_1974, maccrimmon_utility_1979}. Their conclusions inevitably depend on how axioms have been presented to the individuals and on whether counter-arguments have been presented, as \citet{slovic_who_1974} themselves point out. But lacking a systematic approach to determine stability facing counter-arguments, those conclusions only shed very partial light on the question of acceptance of specific norms by the general public. The notion of deliberated preferences can help confront different possible norms for normative decision theory by seeing them as constituents of theories of deliberated preferences and using the procedure described in this article to confront theories.

Second, the concept of nudge \citep{thaler_nudge:_2009} has been tremendously popular in the recent past, but has also been vastly criticized: the risk of manipulation is particularly high whenever it is unsure that the individual would indeed have adopted the decision that the nudge suggests after careful thinking.

Third, analytical philosophy has presented very elaborated and diverging views about the concept of justice. \citet{rawls_theory_1999}, in particular, suggests that the appropriate conception of justice is one that individuals would arrive at in a state of reflexive equilibrium (a concept borrowed from \citet{goodman_fact_1983}). \citeauthor{rawls_theory_1999} and \citeauthor{goodman_fact_1983} insist that individuals must not only accept principles, when thinking about difficult notions such as justice, but also think about consequences of those principles: sometimes, principles that appear intuitively acceptable yield consequences that do not appear intuitively acceptable, in which case something must give. The state of reflexive equilibrium is a state where accepted principles yield acceptable consequences only. The concept of deliberated preferences may be considered as defining that state of reflexive equilibrium, or to be a path towards such a goal (an important difference is that it does not explicitly model the difference between principles and consequences). One way to go beyond philosophers disagreement about the right conception of justice, then, is to ask non-philosophers about what they think is right and so use these laymen as arbiters of the competition between views of justice. Similar empirical studies have started to be conducted recently \citep{gaertner_empirical_2012}. But those studies are conducted by asking individuals only which distribution of goods they consider intuitively right in various circumstances. Thus, they do not try to capture their well-considered judgments for justice, where different reasonings in favor of different conceptions of justice would be presented to the individuals together with the distributions. Hence, this does not test appropriately whether laymen would accept, after thinking carefully, the sophisticated conceptions of justice presented by a given philosopher or economist. A more appropriate way of testing the convincing power of elaborated claims about justice may be considered to require, here again, to present his conception of justice to the individual and submit it to counter-arguments (possibly coming from other sophisticated thinkers).

The first and third contexts of usage come together when it is desired to test whether an individual accepts some axiom as representing an acceptable way of reasoning for him. One simple way is to explain axioms individually and ask the individual whether they appear acceptable. But this fails to make sure the individual understands and accepts the consequences of the union of the axioms accepted by such a procedure \citep{meinard_justification_2018}. (Note that probably nobody is able to mentally deduce all consequences of a union of non trivial axioms.) As an example, in a context of trying to determine the voting rule that best represents someone’s notion of justice, the tested individual could accept the axioms of universal domain, independence of irrelevant alternatives, and pareto dominance, without realizing that he has just accepted dictatorship \citep{arrow_social_1963}.

\section{Decision situation}
The object of study in a given decision situation is the deliberated preferences of a given individual $i$, about a topic $\alltopic$, given a set of arguments $\allargs$, given a certain query protocol, and over a certain time frame.

A decision situation is defined as a triple $(\alltopic, \allargs, \ileadstost)$. It concerns a specific individual $i$, whose deliberated preferences are studied, and whose reactions to arguments is modeled by $\ileadstost$, as will be defined shortly.

The decision situation contains a topic denoted by $\alltopic$, which is a set of propositions $\prop \in \alltopic$ about which we are interested of knowing the deliberated preferences of $i$. Propositions are not described further, but are supposedly understandable by $i$ (for example, they could be sentences in some natural language). Alternatively, in the case of an analyst helping $i$ to make a decision, $\alltopic$ is a set of propositions about which $i$ is interested to know his own deliberated preferences. I assume $\alltopic$ is closed under negation and introduce a symbol $¬$ for negating a proposition: if $\prop \in \alltopic$, then $¬\prop \in \alltopic$, and $¬(¬\prop) = \prop$.

With the decision situation comes also a set of all arguments $\allargs$. This set contains all the arguments that may possibly be considered relevant by $i$ for his decision problem, and possibly more.
Importantly, the notion of deliberated preferences does not require to constrain a priori $\allargs$ to some set of arguments that would fit some precise notion of relevancy, coherence, or even well-formedness. This permits to avoid introducing inadequate normative principles into the deliberated preferences: only the individual $i$ is then considered legitimate to dictate what is a relevant argument. Under that view, $\allargs$ may contain anything that can possibly be considered as an argument by anyone, under the widest possible conception of an argument. The notion of deliberated judgment can however also be applied when considering a restricted set of arguments, for example, the arguments that have been put forward by some specific set of experts when talking about the topic $\alltopic$.

The set $\allargs$ may be infinite or very large: as we will see, thanks to our falsificationist approach, there is no need to be able to explore it entirely. The set of arguments contains at least the empty argument, denoted by $\zar$.

\begin{example}
	Consider a decision problem where a set of alternatives $\allalts$ is given, food products among which $i$ would like to form a deliberated preference (taking into account the effects of food on health, price, pleasure, morality issues related to the production process, and so on). The topic could be defined as $\alltopic$ = $\Set{\prop_\alt, ¬\prop_\alt, \forall \alt \in \allalts}$, where $\prop_\alt$ is the proposition according to which $\alt$ is one of the best alternatives among $\allalts$ (there is no strictly better alternatives among $\allalts$, from $i$’s point of view), and $¬\prop_\alt$ is the proposition according to which $\alt$ is not one of the best alternatives among $\allalts$ (some $\alt' \in \allalts$ is a strictly better alternative, from $i$’s point of view). 
	The set $\allargs$ is defined as the set of all strings, because in this example it is considered acceptable to restrict arguments to those that can take a textual form.
\end{example}

As vNM aptly summarize, “It is clear that every measurement – or rather every claim of measurability – must ultimately be based on some immediate sensation, which possibly cannot and certainly need not be analyzed any futher.” The fundamental element here is the reaction of $i$ to arguments. Given a proposition $\prop$ and two arguments $\ar, \ar' \in \allargs$, a query consists in presenting both arguments to $i$ and observing which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind: is it that $\prop$ holds, that $¬\prop$ holds, or that both are possible? The third possibility allows for the case where no argument appears more decisive than the other one.

It is also necessary to capture the evolution of the position of $i$ towards arguments over time: a repeated query using a given pair of arguments may yield different answers, for example because the repeated question has been interleaved with another query containing other arguments (which $i$ may still have in mind when answering the second repetition), or for any other known or unknown reason. The important information for us is the \emph{set} of answers that $i$ could give to a query involving two given arguments and a given topic, among the set $\{\prop, ¬\prop, \text{both}\}$.
Hence, given $\prop \in \alltopic$, define a relation $\ileadsto^\prop \subseteq (\allargs × \allargs) × \powersetz{\{\prop, ¬\prop, \text{both}\}}$, where $\powersetz{B}$ designates the subsets of $B$ excluding the emptyset, with the semantics that $(\ar, \ar') \ileadsto^\prop C$, with $C \subseteq \Set{\prop, ¬\prop, \text{both}}$, iff $\forall c \in C$, it is observable at least once that $i$, when presented with $(\ar, \ar')$, considers $c$ valid in her current state of mind. The querying protocol is supposedly made so that for any $\prop \in \alltopic$, $\ileadsto^\prop = \ileadsto^{¬\prop}$.

This information is only indirectly observable: for example, if $(\ar, \ar') \ileadsto^\prop \{\prop\}$, a querier will never know more than $(\ar, \ar') \ileadsto^\prop C$ with $\prop \in C \subseteq \Set{\prop, ¬\prop, \text{both}}$. This limitation of our observation will not be a problem for the results of this article. Intuitively, this bears on the fact that if a theory claims that $(\ar, \ar') \ileadsto^\prop \{\prop\}$, then either this claim is correct, or it is falsifiable, in the sense that there is a possibility of observing a contradiction to its claim.

I voluntarily leave the querying protocol not precisely specified, as multiple reasonable choices are possible and the results of this article depend only on specific assumptions that will be made clear shortly. One possible choice, when arguments and propositions are strings, is to read both arguments to $i$, together with $\prop$, and ask $i$ to verbally report the choice he opts for among the three possibilities. Another is to present both arguments and ask $i$ to pick an item among some choice set and make the choice correspond to a validation of $\prop$ or $¬\prop$, possibly even letting the choice engage $i$, for example telling $i$ that she may keep the object of choice (this is illustrated in \cref{ex:pick}). This second possibility is especially interesting as it makes the protocol partially observable in the restricted sense usually appreciated in revealed preferences approaches. Note that depending on the protocol, the choice “both” is not necessarily observable (more about this below).
%In that case, the semantics of $\prop$ and $¬\prop$ relatedly change: instead of meaning that $\prop$ is definitely the 
%Indeed, \citet{tversky_intransitivity_1969} note that individuals “are not perfectly consistent in their choices. When faced with repeated choices between x and y, people often choose x in some instances and y in others. Furthermore, such inconsistencies are observed even in the absence of systematic changes in the decision maker’s taste which might be due to learning or sequential effects. It seems, therefore, that the observed inconsistencies reflect inherent variability or momentary fluctuation in the evaluative process.” This is an accepted fact of experimental psychology \citep{luce_utility_2000}. However, it is reasonable to suspect that fluctuations occur less often, if at all, for some choices. For example, facing a choice between tea or coffee, I would virtually never pick tea (all other things being equal). We are interested in capturing precisely those kind of preferences.

One can also use $\zar$ as one or both of the arguments, in which case the protocol queries the preference of $i$ given only one or given no argument.

\begin{example}[cont.]
	\label{ex:pick}
	Continuing the example, define the querying protocol as, given $\prop_\alt$, presenting two arguments to $i$ and let $i$ choose an item among $\allalts$. $i$ may keep the chosen item. 
	If $i$ chooses $\alt$, it is considered that $i$ has validated $\prop_\alt$, otherwise, $¬\prop_\alt$. The option “both” is not observable.
%	Queries are separated by one week at least, which (as supposed in this example) is enough to ensure that no effect of pleasure for variability enter into play. 
\end{example}

To summarize this information, define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$, and define a single relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$, pronounced “always leads to”. Elements in $\Phi$ will also be called propositions, which should create no ambiguity. Define $\forall \phi = (\prop, \text{possible}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadsto^\prop C$ with $C \subseteq \Set{\prop, \text{both}}$; and $\forall \phi = (\prop, \text{sure}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$. 
The semantics are that whenever $(\ar, \ar') \ileadstost (\prop, \text{sure})$, $i$ always considers that $\prop$ is sure (meaning that $¬\prop$ is excluded), when presented with these two arguments, and whenever $(\ar, \ar') \ileadstost (\prop, \text{possible})$, $i$ always considers that $\prop$ is “at least” possible (meaning that $¬\prop$ may be possible as well), when presented with these two arguments.

Define the negation operator $¬$ over $\Phi$ as: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. It follows that $¬¬\phi = \phi$. Write $(\ar, \ar') \nileadstost \phi$ for $¬[(\ar, \ar') \ileadstost \phi]$ and define $(\ar, \ar') \ileadstoe \phi$ as equivalent to $(\ar, \ar') \nileadstost ¬\phi$. 
The relation $\ileadstoe$ is pronounced “sometimes leads to”.
A proposition $\phi$ is about $\prop$ iff $\phi \in \{\prop, ¬\prop\} × \{\text{possible}, \text{sure}\}$. 
\footnote{\label{ft:sttoe}From (A1) and (A2) we can write, using this notation, $(\ar, \ar') \ileadstost \phi ⇒ (\ar, \ar') \ileadstoe \phi$.}

Define $f_\prop$ from any unordered pair of arguments $(\ar, \ar')$ into $\powerset{\prop\text{-sure}, ¬\prop\text{-sure}, b)}$ (where $b$ means both possible). We The set designates all cumulated answers over time, thus allowing $i$ to be unstable (because she has learnt new things in between two questioning, or just because of time passing, or for any known or unknown reason). We assume $b$ is observable as well.

@@

When “both” is not observable, the same definitions hold, with $f_t(\ar, \ar') ≠ b$. It follows that $\ileadstopst = \ileadstosst$. We say in this case that $\ileadstost$ does not distinguish sure from possible. Thus, elements $\phi \in \Phi$ may be considered as simply equal to $\prop$ or $¬\prop$, as $(\prop, \text{sure})$ is treated exactly as $(\prop, \text{possible})$.

@@

The following proposition follows from the fact that when $\ar^+$ is $\phidec$, given any $\ar^−$, $(\ar^+, \ar^−) \ileadstost \phi$ hence $(\ar^+, \ar^−) \ileadstoe \phi$ (using \cref{ft:sttoe}).
\begin{proposition}
	\label{thm:nocontrdec}
	$\exists \ar \ \phidec ⇒ \nexists \ar' \ ¬\phidec.$
\end{proposition}

The deliberated judgment of $i$ maps any unordered pair $(\prop, ¬\prop)$ to one of four possibilities. For easier treatment, we summarize $i$’s deliberated judgment in a single set $\iPhi \subseteq \Phi$, as follows. (Interpretation is given in parenthesis.)
\begin{itemize}
	\item $(\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \in \iPhi ∧ (¬\prop, \text{possible}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $\exists \ar \ \propsuredec$ ($\prop$ is sure)
	\item $(\prop, \text{possible}) \in \iPhi ∧ (¬\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$ iff $\exists \ar \ \proppossdec$ and $\exists \ar \ ¬\proppossdec$ (both $\prop$ and $¬\prop$ are possible)
	\item $\phi \notin \iPhi$ for all $\phi$ about $\prop$ iff neither $\prop$ is sure, nor $¬\prop$ is sure, nor both are possible ($(\prop, ¬\prop)$ is undecided).
\end{itemize}
\Cref{thm:nocontrdec} shows that the second case is separate from the first.
\footnote{Example of $(\prop, ¬\prop)\text{-undecided}$: $\forall (\ar, \ar'), (\ar, \ar') \ileadstose ¬\prop, (\ar, \ar') \ileadstose \prop$.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}

\begin{proposition}
	a) $\phi \in \iPhi ⇒$ some $\ar$ is $\phidec$. b) $(\prop, \text{possible}) \in \iPhi$ iff $\exists \ar \ \proppossdec ∧ (\exists \ar \ \propsuredec ∨ \exists \ar \ ¬\proppossdec)$. c) $(\prop, \text{sure}) \in \iPhi ⇒ (\prop, \text{possible}) \in \iPhi$. d) $\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$. e) $(\prop, ¬\prop)$ undecided $⇔ \phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$. f) If $\phi \in \iPhi$ for some $\phi$ about $\prop$, then $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$ for all $\phi$ about $\prop$. g) If $\ileadstost$ does not distinguish sure from possible, no pair of proposition is mapped to “both are possible”.
\end{proposition}
\begin{proof}
	a) is a consequence of A2. b) and c) follow from the definition of $\iPhi$. d) is obtained using a), \cref{thm:nocontrdec}, and the contrapositive of a). About e), we must show that $\phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$ implies $\phi' \notin \iPhi$ for all $\phi'$ about $\prop$. Assume $\phi=(l, \text{possible})$ and $¬\phi = (¬l, \text{sure})$ are not in $\iPhi$, with $l = t$ or $l = ¬t$. We need to prove that $(l, \text{sure})$ and $(¬l, \text{possible})$ are not in $\iPhi$ either. We obtain $(l, \text{sure}) \notin \iPhi$ using the contrapositive of c). And $(¬l, \text{possible}) \notin \iPhi$ because $(¬l, \text{possible}) \in \iPhi$ would require another $\phi'$ about $\prop$ to be present in $\iPhi$, as can be checked from the definition of $\iPhi$. f) follows from d) together with the fact that $(\prop, ¬\prop)$ is (by definition) not undecided when $\phi \in \iPhi$ for some $\phi$ about $\prop$, hence, from e), $\phi \in \iPhi ∨ ¬\phi \in \iPhi$. Finally, g) is because in that case, $\exists \ar \ \proppossdec ⇔ \exists \ar \ \propsuredec$, thus using \cref{thm:nocontrdec}, $\exists \ar \ \proppossdec ⇒ \nexists \ar' \ ¬\proppossdec$.
\end{proof}

\section{Theories proposing decisive arguments}
In most situations, there are too many arguments to exhaustively test whether any $\ar$ is a decisive argument for any $\phi$. Furthermore, it is impossible to confirm definitely that $(\ar, \ar') \ileadstost \phi$, whenever $(\ar, \ar') \ileadstost \phi$. We rather propose to introduce theories, and expose them to falsification.

@@

As is usual in science, falsification procedures should not be considered as ways of definitely validating a theory, but rather as ways of confronting theories.
We say that two theories $T, T'$ disagree on $\phi$ (or equivalently, on $¬\phi$) iff $\phi \in \Phi_{d, T}$ and $¬\phi \in \Phi_{d, T'}$ in some situation $d \in D_T ∪ D_{T'}$. We say that $T$ and $T'$ disagree iff they disagree on some $d \in D_T ∪ D_{T'}$ and $\phi \in \Phi$.
To confront a theory $T$ to a disagreeing theory $T'$, suffice to test, for any $\phi$ such that $\phi \in \Phi_T, ¬\phi \in \Phi_{T'}$, whether $(f_T(\phi), f_{T'}(¬\phi)) \ileadstoe \phi$.
Observe that such a test necessarily falsifies one of the theories.

The following proposition shows that this notion of falsification adequately tests the claims of the theory.
\begin{proposition}
	A theory $T$ proposing decisive arguments has no falsification instances iff $\forall d \in D_T, \phi \in \Phi_{d, T}: f_{d, T}(\phi)$ is $\phidec$. Furthermore, if a theory $T$ proposing decisive arguments has no falsification instances, then $\Phi_{d, T} \subseteq \iPhi$, $\forall d \in D_T$.
\end{proposition}

@@

\bibliography{simple}
